---
title: "Government Spending"
author: "Archana Chittoor"
date: "April 21, 2019"
output: html_document
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## R Markdown

```{r budget}
library(tidyverse)

## Load the data
climate_exp <- read_csv('C:/Users/archa/Documents/Archana/DSP_539/assignment_3/climate_spending.csv')
energy_exp <- read_csv('C:/Users/archa/Documents/Archana/DSP_539/assignment_3/energy_spending.csv') 
fed_rd <- read_csv('C:/Users/archa/Documents/Archana/DSP_539/assignment_3/fed_r_d_spending.csv') 

## Define a function which converts the amount spent to millions of dollars

convert_to_millions <- function(dfvector){
  col_in_millions <- round(dfvector/(10^6),digits=3)
  return(col_in_millions)
}

## Convert the amount in some of the columns to millions of dollars

gcc_spending_millions <- convert_to_millions(climate_exp$gcc_spending)
energy_spending_millions <- convert_to_millions(energy_exp$energy_spending)
fed_rd_millions <- convert_to_millions(fed_rd$rd_budget)

```


## Including Plots

The below plot depicts the amount spent (in millions of dollars) as part of climate spending, over a period of 18 years

```{r, echo=FALSE}

# Order the legend
ordered_legend_clim <- filter(climate_exp,year == max(year)) %>% arrange(desc(gcc_spending)) %>%
  pull(department)
ggplot(climate_exp,aes(x = year, gcc_spending_millions, color=department)) +
  geom_line(size = 1) + scale_color_discrete(breaks = ordered_legend_clim, name = "Department") +
  labs(title = "Climate Spending over the years",y = "Gcc Spending (in millions)", x  = "Year") +
  scale_x_continuous(breaks = seq(min(climate_exp$year),max(climate_exp$year),5))
```

The below plot shows the climate spending for each department in a separate subplot

```{r, echo=FALSE}

ggplot(climate_exp,aes(x = year, y = gcc_spending_millions, color = department)) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + facet_wrap(~department) +
  theme(legend.position = "none") +
  labs(title = "Climate Spending per department",y = "Gcc Spending (in millions)", x  = "Year") +
  scale_x_continuous(breaks = seq(min(climate_exp$year),max(climate_exp$year),5))
```

The below plot indicates the amount spent (in millions of dollars) as part of energy spending, over a period of 22 years

```{r, echo = FALSE}

ordered_legend_energy <- filter(energy_exp,year == max(year)) %>% arrange(desc(energy_spending)) %>%
  pull(department)
ggplot(energy_exp,aes(x = year, energy_spending_millions, color=department)) +
  geom_line(size = 1) + scale_color_discrete(breaks = ordered_legend_energy, name = "Department") +
  labs(title = "Energy spending over the years",y = "Energy Spending (in millions)", x  = "Year") +
  scale_x_continuous(breaks = seq(min(energy_exp$year),max(energy_exp$year),5))
```

This is a plot that shows the energy spent for each department in a separate subplot

```{r, echo= FALSE}
ggplot(energy_exp,aes(x = year, y = energy_spending_millions, color = department)) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + facet_wrap(~department) +
  theme(legend.position = "none") +
  labs(title = "Energy Spending per department",y = "Gcc Spending (in millions)", x  = "Year") +
  scale_x_continuous(breaks = seq(min(energy_exp$year),max(energy_exp$year),5))
```


Plot that indicates the R and D budget (in millions of dollars) over a period of 42 years

```{r, echo = FALSE}
ordered_legend_fedrd <- filter(fed_rd,year == max(year)) %>% arrange(desc(rd_budget)) %>%
  pull(department)
ggplot(fed_rd,aes(x = year, y = fed_rd_millions, color=department)) +
  geom_line(size = 1) + scale_color_discrete(breaks = ordered_legend_fedrd, name = "Department") +
  labs(title = "Federal R and D Budget",y = "R and D Budget (in millions)", x  = "Year") + 
  scale_x_continuous(breaks = seq(min(fed_rd$year),max(fed_rd$year),5))
```


Plot to verify whether the rd_budget is related to gdp every year. There does not seem to be any direct relation between the two

```{r, echo=FALSE}
ggplot(fed_rd, aes(gdp, rd_budget,color = department))+ geom_line(size =1) +
  labs(title = "GDP vs R and D Budget",y = "R and D Budget", x  = "GDP") 

```


Plot to show the gdp increase over the years. As seen in the plot, the gdp has increased linearly over time

```{r, echo=FALSE}
ggplot(fed_rd, aes(year,gdp)) + geom_line(size=1) + 
  scale_x_continuous(breaks = seq(min(fed_rd$year),max(fed_rd$year),5)) +
  labs(title = "GDP over the years",y = "GDP", x  = "Year") 
  
```